---
title: "X-much in IrE and AusE - Part 5: Statistical Analysis"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much structure in Irish and Australian English. 

This part of the analysis performs the data analysis. The data is processed using *tidyverse* package(s) [@tidyverse] and the statistical analysis uses \chi^2^-tests [see @cochran1952x2] and configural frequency analysis [see @lienert1975cfa] using the *cfa* package [@cfa]. Effect sizes are calculated using the *effectsize* package [@effectsize].

# Session preparation

* install packages

```{r install, eval = F, message = F, warning = F}
install.packages("tidyverse")
install.packages("here")
install.packages("readxl")
install.packages("flextable")
install.packages("quanteda")
install.packages("tidytext")
install.packages("cfa")
install.packages("report")
install.packages("effectsize")
```


* load packages

* set options


```{r xmprep_2_01}
# load packages
library(tidyverse)
library(here)
library(readxl)
library(flextable)
library(quanteda)
library(tidytext)
library(cfa)
library(vcd)
library(effectsize)
# setting options
options(stringsAsFactors = F)
```

# Load data

```{r load, message=F, warning=F}
xmuch <- base::readRDS(file = here::here("data", "xmuch.rda"))
# inspect
head(xmuch, 10)
```

Inspect structure of the data

```{r}
str(xmuch)
```


# Statistical Analysis

## Single vs Multiword

RQ: do ire and aus differ regarding the status of x (single word vs multiword)?

```{r}
phrasetb <- xmuch %>%
  dplyr::filter(Data == "Xmuch") %>%
  dplyr::select(Variety, Status) %>%
  dplyr::group_by(Variety, Status) %>%
  dplyr::summarise(Frequency = n()) %>%
  dplyr::group_by(Variety) %>%
  dplyr::mutate(Total = sum(Frequency)) %>%
  dplyr::ungroup() %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Percent = round(Frequency/Total*100, 2))
# inspect
phrasetb
```



```{r}
phrase_x2 <- phrasetb %>%
  dplyr::select(-Total, -Percent) %>%
  tidyr::spread(Status, Frequency) %>%
  as.matrix()
# add rownames
rownames(phrase_x2) <- phrase_x2[, 1]
phrase_x2 <- phrase_x2[, 2:3]
# convert to numeric
phrase_x2 <- t(apply(phrase_x2, 1, function(x){ 
  x <- as.numeric(x) }))
# add column names
#colnames(phrase_x2) <- names(table(phrasetb$element))
colnames(phrase_x2) <- c("single-word", "multi-word exp.")
# inspect
phrase_x2
```

* perform X2-test [@stats]

```{r}
chisq.test(phrase_x2)
# effect size
effectsize::effectsize(chisq.test(phrase_x2), type = "phi")
```


generate visualization


```{r}
# open window
png(here::here("images", "multiword_assoc.png"), width = 500, height = 450)
# generate plot
assocplot(phrase_x2)
# close window
dev.off() 
assocplot(phrase_x2)
```


## Emotionality

tabulation


```{r}
emo_tab <- xmuch %>%
  dplyr::group_by(Variety, Status, Polarity) %>%
  dplyr::summarise(Freq = n()) %>%
  dplyr::group_by(Variety, Status) %>%
  dplyr::mutate(Total = sum(Freq)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Percent = round(Freq/Total*100, 1),
                Frequency = paste0(Freq, " (", Percent, ")")) %>%
  dplyr::ungroup() %>%
  dplyr::select(-Freq, -Total, -Percent) %>%
  tidyr::spread(Polarity, Frequency)
# inspect
emo_tab
```

Calculate totals

```{r}
# au
austot <- 11+63+20+952+0+63
negau <- round(74/austot*100, 1)
nonau <- round(972/austot*100, 1)
posau <- round(63/austot*100, 1)
# ire
iretot <- 4+37+6+929+0+57
negire <- round(41/iretot*100, 1)
nonire <- round(935/iretot*100, 1)
posire <- round(57/iretot*100, 1)
# results
negau
nonau
posau
negire
nonire
posire

```



```{r}
xmdtb <- xmuch %>%
  dplyr::select(Variety, X, POS, Emotionality, Polarity, Status) %>%
  dplyr::rename(word = X)
# inspect
head(xmdtb)
```


* tabulation

```{r tb}
tb1 <-base::readRDS(file = here::here("tables", "tb1.rda")) %>%
  dplyr::mutate(Variety = stringr::str_replace_all(text, ".*_([a-z]{2,2})_.*", "\\1"),
                Variety = ifelse(Variety == "au", "Australia", "Ireland")) %>%
  dplyr::group_by(Variety) %>%
  dplyr::summarise(words = sum(words),
                   texts = n())
# inspect
head(tb1)
```


```{r}
xmdtb %>%
  dplyr::group_by(Variety, Status) %>%
  dplyr::summarise(Frequency = n())
```

### X2 (wo variety)

RQ: Does the emotionality of words in the test data (in the x-much construction) differ from the emotionality in the control data?

```{r}
emo_tb1 <- xmdtb %>%
  dplyr::group_by(Emotionality, Status) %>%
  dplyr::summarise(Frequency = n())
# inspect
head(emo_tb1)
```

Perform x2-test

```{r}
emo_x2 <- emo_tb1 %>%
  tidyr::spread(Status, Frequency) %>%
  as.matrix()
rn<- emo_x2[,1]
emo_x2 = emo_x2[, 2:3]
emo_x2 <- apply(emo_x2, 2, as.numeric)
rownames(emo_x2) <- rn
emo_x2
# perform x2 test
chisq.test(emo_x2)
# effect size
effectsize::effectsize(chisq.test(emo_x2), type = "phi")
```

The words in the x-much structure are significantly more emotional compared to the words in the control data.

```{r}
# adapt column names
colnames(emo_x2) <- c("control", "X-much")

# open window
png(here::here("images", "emo_assoc.png"), width = 500, height = 350)
# generate plot
assocplot(emo_x2)
# close window
dev.off() 
assocplot(emo_x2)
```

### X2 (vareity)


RQ: Does the emotionality of words in the X-much structure in Australia differ from the emotionality of words in the X-much structure in Ireland?

```{r}
emo_tb2 <- xmuch %>%
  dplyr::filter(Data == "Xmuch") %>%
  dplyr::group_by(Emotionality, Variety) %>%
  dplyr::summarise(Frequency = n())
# inspect
head(emo_tb2)
```


Perform x2-test

```{r}
emo_x22 <- emo_tb2 %>%
  tidyr::spread(Variety, Frequency) %>%
  as.matrix()
rn<- emo_x22[,1]
emo_x22 = emo_x22[, 2:3]
emo_x22 <- apply(emo_x22, 2, as.numeric)
rownames(emo_x22) <- rn
emo_x22
# perform x2 test
chisq.test(emo_x22)
# effect size
effectsize::effectsize(chisq.test(emo_x22), type = "phi")
```


Australia and Ireland differ significantly with respect to the emotionality of words in the x-much structure.

```{r}
# open window
png(here::here("images", "emo_var_assoc.png"), width = 500, height = 350)
# generate plot
assocplot(emo_x22)
# close window
dev.off() 
assocplot(emo_x22)
```

###  CFA

```{r}
emo_tb3 <- xmdtb %>%
  dplyr::group_by(Emotionality, Variety, Data) %>%
  dplyr::summarise(Frequency = n())
# inspect
emo_tb3
```

```{r}
configs <- emo_tb3 %>%
  dplyr::select(Variety, Emotionality, Data)
counts = emo_tb3$Frequency
cfa::cfa(configs, counts)
```

There are significantly more negative adjectives in the x-much construction in the Australian data after corrections compared to any other configuration.

### Visualization

* prepare data

```{r}
emo_vis <- emo_tb3 %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.character, factor) %>%
  dplyr::mutate(Data = ifelse(Data == "Xmuch", "X-much", "control"))
# inspect
head(emo_vis)
```


```{r}
emo1 <- emo_vis %>%
  dplyr::filter(Emotionality == "emotional") %>%
  dplyr::pull()
emo2 <- emo_vis %>%
  dplyr::filter(Emotionality == "nonemotional") %>%
  dplyr::pull()
# add dimnames
column.names <- c("Australia", "Ireland")
row.names <- c("control", "X-much")
matrix.names <- c("emotional", "nonemotional")
# generate matrix
emo_mx <- array(c(emo1, emo2), dim = c(2, 2, 2),
                  dimnames = list(row.names, 
                                  column.names,
                                  matrix.names))
# inspect
emo_mx
```



Generate  mosaic plot


```{r}
# open connection
png(here::here("images", "emo_mosaic.png"), width = 750, height = 300)
# generate plot
mosaic(emo_mx,
       axis.cex = 15,
       shade = TRUE,
       direction = c("h",  "v", "v"),
       just_labels = c("center", "center", "center", "center"))
# close window
dev.off() 
# show plot
mosaic(emo_mx,
       shade = TRUE,
       direction = c("h",  "v", "v"),
       just_labels = c("center", "center", "center", "center"))
```

## Polarity

### X2 (wo variety)

RQ: Does the polarity of words in the x-much structure differ from the polarity in the control data?

Prepare data

```{r}
pol_tb1 <- xmuch %>%
  dplyr::group_by(Polarity, Status) %>%
  dplyr::summarise(Frequency = n())
# inspect
pol_tb1
```

Perform x2-test

```{r}
pol_x2 <- pol_tb1 %>%
  tidyr::spread(Status, Frequency) %>%
  as.matrix()
# inspect
pol_x2
```


```{r}
rn <- pol_x2[,1]
pol_x2 = pol_x2[, 2:3]
pol_x2 <- apply(pol_x2, 2, as.numeric)
rownames(pol_x2) <- rn
pol_x2
# perform x2 test
chisq.test(pol_x2)
# effect size
effectsize::effectsize(chisq.test(pol_x2), type = "phi")
```

The polarity of words in the X-much structure is significantly different from the polarity of words in the control data.

```{r}
# adapt column names
colnames(pol_x2) <- c("control", "X-much")

# open window
png(here::here("images", "pol_assoc.png"), width = 500, height = 350)
# generate plot
assocplot(pol_x2)
# close window
dev.off() 
assocplot(pol_x2)
```


### X2 (wo variety)

RQ: Does the polarity of words in the x-much structure in Australia differ from the polarity of words in the x-much structure in  Ireland?

Prepare data

```{r}
pol_tb2 <- xmuch %>%
  dplyr::filter(Status == "Xmuch") %>%
  dplyr::group_by(Polarity, Variety) %>%
  dplyr::summarise(Frequency = n())
# inspect
pol_tb2
```

Perform x2-test

```{r}
pol_x22 <- pol_tb2 %>%
  tidyr::spread(Variety, Frequency) %>%
  as.matrix()
# inspect
pol_x22
```


```{r}
rn <- pol_x22[,1]
pol_x22 = pol_x22[, 2:3]
pol_x22 <- apply(pol_x22, 2, as.numeric)
rownames(pol_x22) <- rn
pol_x22
# perform x2 test
chisq.test(pol_x22)
# effect size
effectsize::effectsize(chisq.test(pol_x22), type = "phi")
```


Australia and Ireland do  **not**  differ significantly with respect to the polarity of the words in the X-much structure.

```{r}
# open window
png(here::here("images", "pol_var_assoc.png"), width = 500, height = 350)
# generate plot
assocplot(pol_x22)
# close window
dev.off() 
assocplot(pol_x22)
```

### CFA

```{r}
pol_tb1 <- xmuch %>%
  dplyr::group_by(Polarity, Variety, Status) %>%
  dplyr::summarise(Frequency = n())
# inspect
pol_tb1
```

```{r}
configs <- pol_tb1 %>%
  dplyr::select(Variety, Polarity, Status)
counts = pol_tb1$Frequency
cfa::cfa(configs, counts)
```

### Visualization

* prepare data

```{r}
pol_vis <- pol_tb1 %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.character, factor)
# inspect
pol_vis
```


* convert to matrix


```{r}
pol1 <- pol_vis %>%
  dplyr::filter(Polarity == "positive") %>%
  dplyr::pull(Frequency)
pol2 <- pol_vis %>%
  dplyr::filter(Polarity == "neutral") %>%
  dplyr::pull(Frequency)
pol3 <- pol_vis %>%
  dplyr::filter(Polarity == "negative") %>%
  dplyr::pull(Frequency)
# add dimnames
column.names <- c("Australia", "Ireland")
row.names <- c("control", "X-much")
matrix.names <- c("positive", "neutral", "negative")
# generate matrix
pol_mx <- array(c(pol1, pol2, pol3), dim = c(2, 2, 3),
                  dimnames = list(row.names, 
                                  column.names,
                                  matrix.names))
# inspect
pol_mx
```



* mosaic plot

```{r}
# open window
png(here::here("images", "pol_mosaic.png"), width = 750, height = 300)
# generate plot
mosaic(pol_mx,
       shade = TRUE,
       direction = c("h",  "v", "v"),
       just_labels = c("center", "center", "center", "center"))
# close window
dev.off() 
# show plot
mosaic(pol_mx,
       shade = TRUE,
       direction = c("h",  "v", "v"),
       just_labels = c("center", "center", "center", "center"))
```


* alternative mosaic plot

```{r}
# define cols
mycols <- rev(RColorBrewer::brewer.pal(3, "Greys"))
vcd::mosaic(Polarity ~ Variety, data = pol_vis, 
            direction = c("v",  "h"),
            rot_labels = c(0, 0, 0, 0),
            highlighting_fill = mycols,
            labeling_args = list(tl_labels = c(F, T),
                                 set_varnames = c(Polarity = "Polarity\n\n\n\n\n\n\n"),
                                 offset_varnames = c(0, 0, 0, 1),
                                 just_labels = c("center", "center", "center", "right")))
# save plot
#ggsave(here::here("images", "pol_var.png"), units = "cm", width = 8, height = 6)
```



## POS

tabulation


```{r}
pos_tab <- xmuch %>%
  dplyr::group_by(Variety, Status, pos) %>%
  dplyr::summarise(Frequency = n()) %>%
  dplyr::group_by(Variety, Status) %>%
  dplyr::mutate(Total = sum(Frequency)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Percent = round(Frequency/Total*100, 1),
                Frequency = paste0(Frequency, " (", Percent, ")")) %>%
  dplyr::ungroup() %>%
  dplyr::select(-Frequency, -Total, -Percent) %>%
  tidyr::spread(pos, Frequency)
# inspect
pos_tab
```


RQ: Do the word classes (parts-of-speech) of words in the x-much structure differ from the word classes in the control data?


### X2 (wo variety)

Prepare data

```{r}
pos_tb1 <- xmuch %>%
  dplyr::group_by(POS, Status) %>%
  dplyr::summarise(Frequency = n())
# inspect
pos_tb1
```

Perform x2-test

```{r}
pos_x2 <- pos_tb1 %>%
  tidyr::spread(Status, Frequency) %>%
  as.matrix()
# inspect
pos_x2
```


```{r}
rn <- pos_x2[,1]
pos_x2 = pos_x2[, 2:3]
pos_x2 <- apply(pos_x2, 2, as.numeric)
rownames(pos_x2) <- rn
pos_x2
# perform x2 test
chisq.test(pos_x2)
# effect size
effectsize::effectsize(chisq.test(pos_x2), type = "phi")
```

The word classes (parts-of-speech) of words in the x-much structure differ significantly from the word classes in the control data!

generate association plot

```{r}
# adapt column names
colnames(pos_x2) <- c("control", "X-much")

# open window
png(here::here("images", "pos_assoc.png"), width = 500, height = 350)
# generate plot
assocplot(pos_x2)
# close window
dev.off() 
assocplot(pos_x2)
```

### X2 (variety)

RQ: Do Australia and Ireland differ with respect to the word classes (parts-of-speech) of words in the x-much structure?

Prepare data

```{r}
pos_tb2 <- xmuch %>%
  dplyr::filter(Status == "Xmuch") %>%
  dplyr::group_by(POS, Variety) %>%
  dplyr::summarise(Frequency = n())
# inspect
pos_tb2
```

Perform x2-test

```{r}
pos_x22 <- pos_tb2 %>%
  tidyr::spread(Variety, Frequency) %>%
  as.matrix()
# inspect
pos_x22
```


```{r}
rn <- pos_x22[,1]
pos_x22 = pos_x22[, 2:3]
pos_x22 <- apply(pos_x22, 2, as.numeric)
rownames(pos_x22) <- rn
pos_x22
# perform x2 test
chisq.test(pos_x22)
# effect size
effectsize::effectsize(chisq.test(pos_x22), type = "phi")
```

Australia and Ireland differ significantly with respect to the word classes (parts-of-speech) of words in the x-much structure!

generate association plot

```{r}
# open window
png(here::here("images", "pos_var_assoc.png"), width = 500, height = 350)
# generate plot
assocplot(pos_x22)
# close window
dev.off() 
assocplot(pos_x22)
```

### CFA

process data

```{r}
pos_tb2 <- xmuch %>%
  dplyr::group_by(POS, Variety, Status) %>%
  dplyr::summarise(Frequency = n())
# inspect
pos_tb2
```

perform CFA

```{r}
configs <- pos_tb2 %>%
  dplyr::select(Variety, POS, Status)
counts = pos_tb2$Frequency
cfa::cfa(configs, counts)
```


### Visualization

* prepare data

```{r}
pos_vis <- pos_tb2 %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.character, factor)
# inspect
pos_vis
```



* prepare data


```{r}
pos1 <- pos_vis %>%
  dplyr::filter(POS == "Adjective") %>%
  dplyr::pull(Frequency)
pos2 <- pos_vis %>%
  dplyr::filter(POS == "Noun") %>%
  dplyr::pull(Frequency)
pos3 <- pos_vis %>%
  dplyr::filter(POS == "other") %>%
  dplyr::pull(Frequency)
pos4 <- pos_vis %>%
  dplyr::filter(POS == "Verb") %>%
  dplyr::pull(Frequency)
# add dimnames
column.names <- c("Australia", "Ireland")
row.names <- c("control", "X-much")
matrix.names <- c("Adj.", "Noun", "other", "Verb")
# generate matrix
pos_mx <- array(c(pol1, pol2, pol3), dim = c(2, 2, 4),
                  dimnames = list(row.names, 
                                  column.names,
                                  matrix.names))
# inspect
pos_mx
```

Generate mosaic plot

```{r}
# open window
png(here::here("images", "pos_mosaic.png"), width = 750, height = 300)
# generate plot
mosaic(pos_mx,
       shade = TRUE,
       direction = c("h",  "v", "v"),
       just_labels = c("center", "center", "center", "center"),
       rot_labels=c(0,0,90,90),
       offset_labels = c(0, 0, 0.5,0))
# close window
dev.off() 
# show plot
mosaic(pos_mx,
       shade = TRUE,
       direction = c("h",  "v", "v"),
       just_labels = c("center", "center", "center", "center"),
       rot_labels=c(0,0,90,90),
       offset_labels = c(0, 0, 0.5,0))
```



## Words


RQ: Are there Words that are significantly over-represented in the X-much structure?


**Constructionalization more advanced in AusE**

### CFA (wo vareity)

RQ: Regardless of variety, are there words that are significantly attracted by the X-much structure?

```{r}
wordcat_tb <- xmuch %>%
  dplyr::group_by(Status, X) %>%
  dplyr::mutate(Frequency = n()) %>%
  dplyr::mutate(x = ifelse(Frequency < 2, "other", x)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(x, Status) %>%
  dplyr::summarise(Frequency = n())
configs <- wordcat_tb %>%
  dplyr::select(Status, X)
counts = wordcat_tb$Frequency
cfa::cfa(configs, counts)

```

```{r}
r7 <- cfa::cfa(configs, counts)
# save to disc
sig_words <- r7$table %>%
  as.data.frame() %>%
  dplyr::mutate(expected = round(expected, 1),
                Q = round(Q, 3),
                chisq = round(chisq, 3),
                z = round(z, 3),
                p.z = round(p.z, 3),
                label = stringr::str_remove_all(label, "Xmuch ")) %>%
  dplyr::select(-p.chisq, -sig.chisq)
# save to disc
write.table(sig_words, 
            here::here("tables", "xmuch_words.txt"), 
            sep = "\t", 
            row.names = F)
# extract words
words <- r7$table %>%
  as.data.frame() %>%
  dplyr::mutate(Type = ifelse(n > expected, "Type", "Antitype")) %>%
  dplyr::filter(sig.z == T,
                Type == "Type") %>%
  dplyr::pull(label) %>%
  stringr::str_remove_all("Xmuch ")
words
```

Words that are significantly over-represented in the X-much structure: `r words`

### X2 (w variety)

RQ: Do the words that are significantly attracted by the x-much construction differ across varieties?

```{r}
wordcat_tb <- xmuch %>%
  dplyr::filter(Status == "Xmuch") %>%
  dplyr::group_by(Variety, x) %>%
  dplyr::mutate(Frequency = n()) %>%
  dplyr::mutate(x = ifelse(Frequency < 2, "other", x)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(X, Variety) %>%
  dplyr::summarise(Frequency = n())
configs <- wordcat_tb %>%
  dplyr::select(Variety, X)
counts = wordcat_tb$Frequency
cfa::cfa(configs, counts)
```

```{r}
r8 <- cfa::cfa(configs, counts)
words <- r8$table %>%
  as.data.frame() %>%
  dplyr::filter(sig.z == T) %>%
  dplyr::pull(label) %>%
  stringr::str_remove_all("Xmuch ")
words
```

The analysis does not confirm variety specific attraction of types to the X-much structure.


### CFA (w variety)

RQ: Do the words that are significantly attracted by the x-much construction differ across varieties when we consider the control data?

```{r}
wordcat_tb <- xmuch %>%
  dplyr::group_by(Status, Variety, X) %>%
  dplyr::mutate(Frequency = n()) %>%
  dplyr::mutate(X = ifelse(Frequency < 2, "other", X)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(X, Variety, Status) %>%
  dplyr::summarise(Frequency = n())
configs <- wordcat_tb %>%
  dplyr::select(Status, Variety, X)
counts = wordcat_tb$Frequency
cfa::cfa(configs, counts)
```

```{r}
r8 <- cfa::cfa(configs, counts)
# save to disc
config_words <- r8$table %>%
  as.data.frame() %>%
  dplyr::mutate(expected = round(expected, 1),
                Q = round(Q, 3),
                chisq = round(chisq, 3),
                z = round(z, 3),
                p.z = round(p.z, 3),
                label = stringr::str_remove_all(label, "Xmuch ")) %>%
  dplyr::select(-p.chisq, -sig.chisq)
# save to disc
write.table(config_words, 
            here::here("tables", "xmuch_wordconfigs.txt"), 
            sep = "\t", 
            row.names = F)
# save to disc
words <- r8$table %>%
  as.data.frame() %>%
  dplyr::filter(stringr::str_detect(label, "control", negate = TRUE),
                sig.z == T) %>%
  dplyr::pull(label) %>%
  stringr::str_remove_all("Xmuch ")
words
```

The words that are attracted to the X-much structure in Ireland are less negative compared to the words that are attracted to the x-much construction in Oz. Compare *excuses*, *coincidence*, or *agenda* (Ireland) to *jealous*, *paranoid*, *hypocrite*, *creepy*, *threatened*, or *entitled* (Australia).


# Outro

```{r}
sessionInfo()
```


# References


