---
title: "X-much in IrE and AusE - Part 4: Data Processing"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much construction in Irish and Australian English. 

This part of the analysis performs the data processing of the control corpus and combines the controll with the test data.

# Session preparation

* install packages

```{r install, eval = F, message = F, warning = F}
install.packages("tidyverse")
install.packages("here")
install.packages("quanteda")
install.packages("tidytext")
```


* load packages

* set options


```{r xmprep_2_01}
# load packages
library(tidyverse)
library(here)
library(quanteda)
library(tidytext)
# setting options
options(stringsAsFactors = F)
```

# Load data

```{r load, message=F, warning=F}
# load raw control
ccps  <- base::readRDS(file = here::here("data", "ccorpus.rda")) %>%
  dplyr::mutate(docname = stringr::str_remove_all(paste0(file, "_", subfile), "/"))
# load raw xmuch
xmd <- base::readRDS(file = here::here("data/kwics", "xmd.rda")) %>%
  dplyr::mutate(control = "test") %>%
  dplyr::rename(pos = type,
                kwic = clean_content) %>%
  dplyr::select(-pre, -keyword, -post, -PrePosTag, -PostPosTag)
xmdfiles <- unique(stringr::str_remove_all(xmd$docname, "\\..*"))
# inspect
head(xmd, 10)
```

* remove files that are in the xmd data from the controll data

```{r filter, message=F, warning=F}
ccps <- ccps %>%
  dplyr::filter(!docname %in% xmdfiles) %>%
  dplyr::mutate(variety = stringr::str_replace_all(docname, ".*_([a-z]{2,2})_.*", "\\1"))
# inspect
nrow(ccps)
```

inspect data

```{r}
# inspect
head(ccps, 10)
```


* tokenize control data so that every row represents one word

* add variety column

* extract and simplify parts-of-speech

* select a sample of 1000 words (that do not contain special characters)

```{r adj, message=F, warning=F}
# oz control
cctxt_au <- ccps %>%
  dplyr::filter(!docname %in% xmdfiles) %>%
  dplyr::mutate(text = paste0(variety, "_", text)) %>%
  dplyr::filter(stringr::str_detect(text, "^au")) %>%
  dplyr::pull(text)
# ie control
cctxt_ie <- ccps %>%
  dplyr::filter(!docname %in% xmdfiles) %>%
  dplyr::mutate(text = paste0(variety, "_", text)) %>%
  dplyr::filter(stringr::str_detect(text, "^ie")) %>%
  dplyr::pull(text)
# tok oz
tokcc_au <- quanteda::tokens(cctxt_au, what = "fastestword") %>%
  unlist() %>%
  sample(3000)
# tok ie
tokcc_ie <- quanteda::tokens(cctxt_ie, what = "fastestword") %>%
  unlist() %>%
  sample(3000)
# combine
tokcc <-c(tokcc_au, tokcc_ie)
# generate df
tkdf <- tokcc %>%
  as.data.frame() %>%
  dplyr::mutate(variety = c(rep("au", length(tokcc_au)), rep("ie", length(tokcc_ie)))) %>%
  dplyr::mutate(pos = stringr::str_remove_all(tokcc, ".*/"),
                word = tolower(stringr::str_remove_all(tokcc, "/.*")),
                variety = ifelse(variety == "au", "Australia", "Ireland")) %>%
  dplyr::rename(full = colnames(.)[1]) %>%
  dplyr::mutate(pos = stringr::str_remove_all(pos, "[0-9].*|@.*|_.*")) %>%
  dplyr::filter(!stringr::str_detect(word, "\\W")) %>%
  dplyr::group_by(variety) %>%
  dplyr::sample_n(1000)
# inspect
head(tkdf); table(tkdf$variety)
```

# Sentiment analysis

* extract emotives

```{r emo1, message=F, warning=F}
# sentiments
emo <- get_sentiments("nrc") %>%
  dplyr::filter(sentiment == "positive"| sentiment == "negative")
wordpos <- emo %>%
  dplyr::filter(sentiment == "positive") %>%
  dplyr::pull(word)
wordneg <- emo %>%
  dplyr::filter(sentiment == "negative") %>%
  dplyr::pull(word)
```

* annotate emotionality

```{r emo1, message=F, warning=F}
cemodf <- tkdf %>%
  dplyr::mutate(emo = dplyr::case_when(word %in% wordpos ~ "positive",
                                       word %in% wordneg ~ "negative",
                                       TRUE ~ "nonemotional")) %>%
  dplyr::mutate(control = "control") %>%
  dplyr::rename(x = word) %>%
  dplyr::select(-full)
# inspect
head(cemodf)
```

# combine x-much and control

* inspect data

```{r}
head(xmd, 3)
head(cemodf, 3)
```

* combine data (add rows; warning: column names different, hence dplyr::bind_rows)

```{r}
xmuch <- dplyr::bind_rows(xmd, cemodf) %>%
  dplyr::mutate(element = ifelse(is.na(element), "word", element),
                polarity = emo,
                emo = ifelse(emo == "nonemotional", "nonemotional", "emotional"),
                pos = dplyr::case_when(stringr::str_detect(pos, "vvg") ~ "Noun",
                                       stringr::str_detect(pos, "vvn") ~ "Noun",
                                       stringr::str_detect(pos, "nn") ~ "Noun",
                                       stringr::str_detect(pos, "np") ~ "Noun",
                                       stringr::str_detect(pos, "vv") ~ "Verb",
                                       stringr::str_detect(pos, "jj") ~ "Adjective",
                                       TRUE ~ "other")) %>%
  dplyr::rename()
# inspect
head(xmuch)
```

* check pos

```{r}
table(xmuch$pos)
```



# Save data to disc{-}

```{r}
base::saveRDS(xmuch, file = here::here("data", "xmuch.rda"))
```


# Outro

```{r}
sessionInfo()
```





