---
title: "X-much CXN in IrE - Part 2: Data processing"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much construction in Irish English. 

This part of the analysis performs the data processing and extracting examples of the X-much construction.

# Session preparation

cleaning the workspace, loading packages, setting options, and defining paths.

```{r xmprep_2_01}
# load packages
library(tidyverse)
library(here)
library(readxl)
library(flextable)
library(quanteda)
# setting options
options(stringsAsFactors = F)
# set paths
cpath <- "D:\\Uni\\Korpora\\Edited\\GloWbe_postagged"
```

# Extract and process texts from corpus


```{r}
cfiles <- list.files(here::here("data/GloWbe_pos"), pattern = ".rda", full.names = T)
cfiles <- cfiles[stringr::str_detect(cfiles, "_ie_|_au_")]
# inspect
cfiles
```

load texts

```{r}
# load texts
texts <- sapply(cfiles, function(x){
  filename <- stringr::str_replace_all(x, ".*/(.*?).rda", "\\1")
  x <- base::readRDS(x, "rb")
  x <- paste0(x, collapse = " ")
  x <- str_split(stringr::str_replace_all(x, "##", "~~~##"), "~~~")
  x <- unlist(x)
})
# inspect data
str(texts)
```

Clean data

```{r xm_2_05, eval=T}
# simplify names
names(texts) <- stringr::str_replace_all(names(texts), ".*/(.*?).rda", "\\1")
# check number of elements with each file
idx <- sapply(texts, function(x) x <- length(x))
# repeat file names as many times as there are elements in each file.
files <- rep(names(texts), idx)
# convert into data frame
textdf <- texts %>%
  unlist() %>%
  as.vector() %>%
  as.data.frame() %>%
  dplyr::mutate(file = files) %>%
  dplyr::rename(doc_id = colnames(.)[1],
                text = colnames(.)[1]) %>%
  dplyr::mutate(subfile = stringr::str_extract(text, "##.*?\\W"),
                id = 1:nrow(.)) %>%
  tidyr::drop_na()
# remove texts from memory
texts <- NULL
# inspect
head(textdf)
```


# Create control corpus

```{r}
ccorpus <- textdf %>%
  sample_n(1000)
# save to disc
base::saveRDS(ccorpus, file = here::here("data", "ccorpus.rda"))
# inpsect
nrow(ccorpus)
```


# Create test corpus


```{r xm_2_05, eval=T, meassage = F, warning=F}
textc <- textdf %>%
  quanteda::corpus()
# extract names
nms <- stringr::str_remove_all(paste0(textdf$file, "_", textdf$subfile), "/")
# save textdf
base::saveRDS(textdf, file = here::here("data", "textdf.rda"))
# remove texts from memory
textdf <- NULL
# tokenize text corpus
toks <- quanteda::tokens(textc, what = "fastestword")
# add names 
names(toks) <- nms
# inspect
str(toks)
```

## Table with words per document

```{r tb1, message=F, warning=F}
words <- sapply(toks, function(x){
  x <- stringr::str_remove_all(x, "<.*?>")
  x <- stringr::str_remove_all(x, "\\W")
  x <- x[x != ""]
  x <- length(x)
})
tb1 <- data.frame(names(words), words) %>%
  dplyr::rename(text = colnames(.)[1])
# save to disc
base::saveRDS(tb1, file = here::here("tables", "tb1.rda"))
# inspect
head(tb1)
```


## Concordancing of *much*

Split toks and concordancing

```{r}
# split
toks01 <- toks[1:25000]
# kwic
kwic_much_01 <- quanteda::kwic(toks01, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_01, file = here::here("data/kwics", "kwic_much_01.rda"))
# remove texts from memory
toks01 <- NULL
kwic_much_01 <- NULL
```

```{r}
# split
toks02 <- toks[25001:50000]
# kwic
kwic_much_02 <- quanteda::kwic(toks02, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_02, file = here::here("data/kwics", "kwic_much_02.rda"))
# remove texts from memory
toks02 <- NULL
kwic_much_02 <- NULL
```

```{r}
# split
toks03 <- toks[50001:75000]
# kwic
kwic_much_03 <- quanteda::kwic(toks03, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_03, file = here::here("data/kwics", "kwic_much_03.rda"))
# remove texts from memory
toks03 <- NULL
kwic_much_03 <- NULL
```

```{r}
# split
toks04 <- toks[75001:100000]
# kwic
kwic_much_04 <- quanteda::kwic(toks04, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_04, file = here::here("data/kwics", "kwic_much_04.rda"))
# remove texts from memory
toks04 <- NULL
kwic_much_04 <- NULL
```

```{r}
# split
toks05 <- toks[100001:125000]
# kwic
kwic_much_05 <- quanteda::kwic(toks05, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_05, file = here::here("data/kwics", "kwic_much_05.rda"))
# remove texts from memory
toks05 <- NULL
kwic_much_05 <- NULL
```

```{r}
# split
toks06 <- toks[125001:150000]
# kwic
kwic_much_06 <- quanteda::kwic(toks06, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_06, file = here::here("data/kwics", "kwic_much_06.rda"))
# remove texts from memory
toks06 <- NULL
kwic_much_06 <- NULL
```

```{r}
 # split
toks07 <- toks[150001:175000]
# kwic
kwic_much_07 <- quanteda::kwic(toks07, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_07, file = here::here("data/kwics", "kwic_much_07.rda"))
# remove texts from memory
toks07 <- NULL
kwic_much_07 <- NULL
```

```{r}
# split 
toks08 <- toks[175001:200000]
# kwic
kwic_much_08 <- quanteda::kwic(toks08, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_08, file = here::here("data/kwics", "kwic_much_08.rda"))
# remove texts from memory
toks08 <- NULL
kwic_much_08 <- NULL
```

```{r}
# split
toks09 <- toks[200001:225000]
# kwic
kwic_much_09 <- quanteda::kwic(toks09, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_09, file = here::here("data/kwics", "kwic_much_09.rda"))
# remove texts from memory
toks09 <- NULL
kwic_much_09 <- NULL
```

```{r}
# split
toks10 <- toks[225001:length(toks)]
# kwic
kwic_much_10 <- quanteda::kwic(toks10, window = 20, pattern = "much",
                       valuetype = "regex", case_insensitive = F) %>%
  as.data.frame()
# save to disc
base::saveRDS(kwic_much_10, file = here::here("data/kwics", "kwic_much_10.rda"))
# remove texts from memory
toks10 <- NULL
kwic_much_10 <- NULL
```

## Extract session information{-}

```{r}
sessionInfo()
```


***

The END

***

