---
title: "X-much CXN in IrE - Part 2: Data processing"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much construction in Irish English. 

This part of the analysis performs the data processing and extracting examples of the X-much construction.

# Session preparation

cleaning the workspace, loading packages, setting options, and defining paths.

```{r xmprep_2_01}
# load packages
library(tidyverse)
library(here)
library(readxl)
library(flextable)
library(quanteda)
# setting options
options(stringsAsFactors = F)
```

## Cleaning function

```{r}

clean_kwic <- function(kwics){
  kwics <- kwics %>%
  dplyr::mutate(variety = stringr::str_replace_all(docname, ".*_([a-z]{2,2})_.*", "\\1"),
                clean_content = paste0(pre, " << ", keyword, " >> ", post)) %>%
  dplyr::mutate(clean_content = stringr::str_remove_all(clean_content, 
                                                        "\\/[@%]{0,1}[:alnum:]{1,4}[@%]{0,1}_{0,1}[@%]{0,1}[:alnum:]{0,4}[@%]{0,1}_{0,1}[@%]{0,1}[:alnum:]{0,4}[@%]{0,1}")) %>%
  dplyr::mutate(clean_content = stringr::str_remove_all(clean_content, "@")) %>%
  dplyr::mutate(clean_content = stringr::str_remove_all(clean_content, "%")) %>%
  dplyr::mutate(clean_content = stringr::str_squish(clean_content)) %>%
    dplyr::select(-to, -from, -pattern) %>%
  dplyr::filter(stringr::str_detect(keyword, "^much/"),
                stringr::str_detect(clean_content, " [:alnum:]{1,2} <<", negate = TRUE),
                stringr::str_detect(clean_content, " [VERYvery]{4,4} <<", negate = TRUE),
                stringr::str_detect(clean_content, " [SOso]{2,} <<", negate = TRUE),
                stringr::str_detect(clean_content, " too <<", negate = TRUE),
                stringr::str_detect(clean_content, " [Pp]retty <<", negate = TRUE),
                stringr::str_detect(clean_content, " [Hh]ow <<", negate = TRUE),
                stringr::str_detect(clean_content, " [Nn]ot {0,1}<<", negate = TRUE),
                stringr::str_detect(clean_content, " bit <<", negate = TRUE),
                stringr::str_detect(clean_content, " [a-z]{0,}n['|o]t [a-z]{1,} <<", negate = TRUE),
                stringr::str_detect(clean_content, " out <<", negate = TRUE),
                stringr::str_detect(clean_content, " have <<", negate = TRUE),
                stringr::str_detect(clean_content, " were <<", negate = TRUE),
                stringr::str_detect(clean_content, " becoming <<", negate = TRUE),
                stringr::str_detect(clean_content, " write <<", negate = TRUE),
                stringr::str_detect(clean_content, " hurt <<", negate = TRUE),
                stringr::str_detect(clean_content, " London <<", negate = TRUE),
                stringr::str_detect(clean_content, " attract <<", negate = TRUE),
                stringr::str_detect(clean_content, " get[ting]{0,4} <<", negate = TRUE),
                stringr::str_detect(clean_content, " that <<", negate = TRUE),
                stringr::str_detect(clean_content, " this .{0,3}<<", negate = TRUE),
                stringr::str_detect(clean_content, " [Aafter]{5,5} <<", negate = TRUE),
                stringr::str_detect(clean_content, " pretty <<", negate = TRUE),
                stringr::str_detect(clean_content, " do[esing]{0,3} <<", negate = TRUE),
                stringr::str_detect(clean_content, " without <<", negate = TRUE),
                stringr::str_detect(clean_content, " ' s <<", negate = TRUE),
                stringr::str_detect(clean_content, " n't <<", negate = TRUE),
                stringr::str_detect(clean_content, "How @ <<", negate = TRUE),
                stringr::str_detect(clean_content, " >> [:alnum:]{1,} ", negate = TRUE),
                stringr::str_detect(clean_content, " >> about ", negate = TRUE),
                stringr::str_detect(clean_content, " >> faster ", negate = TRUE),
                stringr::str_detect(clean_content, " >> confusion ", negate = TRUE),
                stringr::str_detect(clean_content, " >> anyway ", negate = TRUE),
                stringr::str_detect(clean_content, " >> later ", negate = TRUE),
                stringr::str_detect(clean_content, " >> [MOREmore]{4,4} ", negate = TRUE),
                stringr::str_detect(clean_content, " >> [LESSless]{4,4} ", negate = TRUE),
                stringr::str_detect(clean_content, " >> [WORSEworse]{5,5} ", negate = TRUE),
                stringr::str_detect(clean_content, " >>$", negate = TRUE),
                stringr::str_detect(clean_content, " >> [:alnum:]{0,} {0,1}\\?/\\? | >> [:alnum:]{0,} {0,1}\\!/\\! | >> [:alnum:]{0,} {0,1}\\./\\. | >> [:alnum:]{0,} {0,1}\\;/\\; ")
                )
}
```


Load and clean data

## Clean data{-}


```{r}
muchkwic01  <- base::readRDS(file = here::here("data/kwics", "kwic_much_01.rda")) 
mck01 <- clean_kwic(kwics = muchkwic01)
# inspect data
head(mck01)
```

```{r}
muchkwic02  <- base::readRDS(file = here::here("data/kwics", "kwic_much_02.rda")) 
mck02 <- clean_kwic(kwics = muchkwic02)
# inspect data
head(mck02)
```

```{r}
muchkwic03  <- base::readRDS(file = here::here("data/kwics", "kwic_much_03.rda")) 
mck03 <- clean_kwic(kwics = muchkwic03)
# inspect data
head(mck03)
```

```{r}
muchkwic04  <- base::readRDS(file = here::here("data/kwics", "kwic_much_04.rda")) 
mck04 <- clean_kwic(kwics = muchkwic04)
# inspect data
head(mck04)
```


```{r}
muchkwic05  <- base::readRDS(file = here::here("data/kwics", "kwic_much_05.rda")) 
mck05 <- clean_kwic(kwics = muchkwic05)
# inspect data
head(mck05)
```



```{r}
muchkwic06  <- base::readRDS(file = here::here("data/kwics", "kwic_much_06.rda")) 
mck06 <- clean_kwic(kwics = muchkwic06)
# inspect data
head(mck06)
```



```{r}
muchkwic07  <- base::readRDS(file = here::here("data/kwics", "kwic_much_07.rda")) 
mck07 <- clean_kwic(kwics = muchkwic07)
# inspect data
head(mck07)
```


```{r}
muchkwic08  <- base::readRDS(file = here::here("data/kwics", "kwic_much_08.rda")) 
mck08 <- clean_kwic(kwics = muchkwic08)
# inspect data
head(mck08)
```


```{r}
muchkwic09  <- base::readRDS(file = here::here("data/kwics", "kwic_much_09.rda")) 
mck09 <- clean_kwic(kwics = muchkwic09)
# inspect data
head(mck09)
```


```{r}
muchkwic10  <- base::readRDS(file = here::here("data/kwics", "kwic_much_10.rda")) 
mck10 <- clean_kwic(kwics = muchkwic10)
# inspect data
head(mck10)
```


## Combine kwics

```{r}
mck <- rbind(mck01, mck02, mck03, mck04, mck05, mck06, mck07, mck08, mck09, mck10) %>%
  dplyr::mutate(id = 1:nrow(.))
# inspect
head(mck); nrow(mck); table(mck$keyword)
```



## Save data{-}

```{r}
# save data
base::saveRDS(mck, file = here::here("data/kwics", "kwics_semiclean.rda"))
write.table(mck, here::here("data/kwics", "kwics_semiclean.txt"), sep = "\t", row.names = F)
# load kwics
cmck  <- mck
# inspect data
head(cmck); nrow(cmck)
```

# Manual check

results of manula check

```{r}
na <- c("48", "73", "80", "135", "143", "185", "228", "280", "297", "318", 
        "368", "429", "509", "520", "675") 
mx <- c("1", "5", "7", "8", "11", "17", "26", "27", "29", "32", "33", "35", 
        "39", "41", "43", "49", "51", "57", "59", "66", "67", "82", "100", 
        "101", "105", "113", "115", "119", "131", "136", "137", "138", "142", 
        "159", "162", "165", "168", "169", "170", "180", "189", "194", "217", 
        "218", "222", "224", "243", "245", "260", "274", "281", "283", "299", 
        "305", "313", "315", "316", "317", "327", "332", "341", "345", "346", 
        "348", "364", "366", "370", "373", "374", "392", "402", "405", "413", 
        "414", "418", "421", "430", "435", "436", "440", "445", "449", "453", 
        "454", "455", "461", "467", "469", "480", "486", "491", "495", "501", 
        "503", "510", "511", "513", "521", "524", "527", "534", "544", "548", 
        "549", "555", "556", "557", "571", "572", "573", "574", "577", "580", 
        "590", "603", "604", "605", "618", "623", "635", "639", "642", "657", 
        "676", "677", "678", "688", "693", "694", "695", "696", "698", "708", 
        "709", "712", "728", "731", "732", "741", "743", "754", "788", "789", 
        "790", "792", "835", "837", "846") 
```




```{r}
lmck <- cmck %>%
  dplyr::filter(id %in% mx)
# inspect data
head(lmck); nrow(lmck)
```




```{r}
# extract pos tags
clean_kwic2 <- lmck %>%
  dplyr::mutate(PrePosTag = str_remove_all(pre, ".*/"),
                PrePosTag = stringr::str_squish(PrePosTag),
                PostPosTag = str_remove_all(post, "^/"),
                PostPosTag = str_remove_all(PostPosTag, "/.*"),
                PostPosTag = str_remove_all(stringr::str_squish(PostPosTag), ".* ") ) %>%
  dplyr::filter(PostPosTag != ".",
                PostPosTag != ",")
# inspect data
head(clean_kwic2); nrow(clean_kwic2)
```


```{r}
clean_kwic3 <- clean_kwic2 %>%
  dplyr::filter(PostPosTag == "?" | PostPosTag == "!" | PostPosTag == "[:alnum:]{1,}" ) %>%
  dplyr::mutate(clean_content = stringr::str_replace_all(clean_content, " ([:punct:])/[:punct:] ", " \\1 "),
                clean_content = stringr::str_replace_all(clean_content, "<\\w{1,2}>", " "),
                clean_content= stringr::str_squish(clean_content)) 
# inspect data
head(clean_kwic3); nrow(clean_kwic3); clean_kwic3$clean_content
```

## Manual correction{-}

```{r oldhits, eval = F}
hits <- c(599, 5525, 5530,5816, 11582, 13200, 17046, 20991, 21795, 30236, 34972, 
          35020, 35120, 35308, 37854, 39600, 40453, 40876, 41221, 41370, 46497, 
          46498, 47517, 51719, 52752, 53110, 57010, 60037, 70653, 70684, 70843, 
          71039, 87422, 89308, 89535)
kwics <- clean_kwic3 %>%
  dplyr::filter(id %in% hits)
```


## Save data to disc{-}

```{r}
write.table(kwics, here::here("data", "clean_kwics.txt"), sep = "\t") 
```


load data

```{r}
xm <- read.delim(here::here("data", "clean_kwics.txt"), sep = "\t")
# inspect
head(xm)
```





# Outro

```{r}
sessionInfo()
```





