---
title: "X-much CXN in IrE - Part 2: Data processing"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much construction in Irish English. 

This part of the analysis performs the data processing and extracting examples of the X-much construction.

# Session preparation

cleaning the workspace, loading packages, setting options, and defining paths.

```{r xmprep_2_01}
# clean workspace
rm(list=ls(all=T))
# load packages
library(tidyverse)
library(here)
library(readxl)
library(flextable)
library(quanteda)
# setting options
options(stringsAsFactors = F)
# set paths
cpath <- "D:\Uni\Korpora\Edited\GloWbe_postagged"
```

# Extarct examples for manual coding

Extract sample and save it to disc

```{r xm_2_03}
# list files
cfiles <- list.files(cpath, pattern = ".txt", full.names = T)
# load texts
texts <- sapply(cfiles, function(x){
   x <- readLines(x)
   x <- paste0(x, collapse = " ")
})
# inspect data
str(texts)
```

extract instances of much

```{r xm_2_07}
# kwic
kwics <- kwic(texts, ".*{5, 75}jj much/[a-z]{1,3}[0-9]{0,}")
# inspect
str(kwics)
```

clean kwics

```{r xm_1_07}
# kwic
kwics <- kwic
# inspect
head(kwics)
```



***

The END

***

