---
title: "X-much CXN in IrE - Part 1: Pos tagging"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much construction in Irish English. 

This part of the analysis performs the part-of-speech tagging of the data.

# Session preparation

In a first step, we prepare the session by cleaning the workspace, loading packages, setting options, and defining paths.

```{r xmprep_1_01}
# load packages
library(tidyverse)
library(here)
library(readxl)
library(flextable)
library(quanteda)
# setting options
options(stringsAsFactors = F)
```

# Writing function for pos tagging

```{r xmpos_1_03}
# write function
glowbe_post <- function(gfiles){
  sapply(gfiles, function(x){
    # extract names
    filename <- str_remove_all(x, ".*/")
    filename <- str_remove_all(filename, ".txt")
    x <- readLines(con = x, n = -1L, ok = TRUE, warn = TRUE, encoding = "unknown", skipNul = T)
    # split data on \t
    x <- strsplit(x, "\t")
    # select first three elements
    x <- sapply(x, function(y){ y <- y[c(1:3)] })
    # convert into data.frames
    x <- t(x)
    x <- as.data.frame(x)
    # paste words and pos tags together
    x <- apply(x, 1, function(y){
      y <- paste(y[1], "/", y[3], sep = "") })
    # paste words together
    x <- paste(x, collapse = " ")
    # add names
    names(x) <- filename
    writeLines(text = x, con = paste("D:\\Uni\\Korpora\\Edited\\GloWbe_postagged/", names(x),".txt", sep = ""))
    rm(list=ls(all=T))
  })
  }
```


apply function to data

extract session information

```{r xmpos_1_05, echo = T, eval = F}
# defining path
path <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Ireland\\WordLemmaPos"
# extract files
gfiles = list.files(path = path, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
# apply function
glowbe_post(gfiles = gfiles) #(remove # to to activate)
```

```{r xmpos_1_05}
# defining path
pathaue <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Australia\\Word lemma pos"
pathbge <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Bangladesh\\Word lemma pos"
pathcae <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Canada\\Word lemma pos"
pathgbb <- "D:\\Uni\\Korpora\\Original\\GloWbE\\GB-Blog\\Word lemma pos"
pathgbg <- "D:\\Uni\\Korpora\\Original\\GloWbE\\GB-General\\Word lemma pos"
pathghe <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Ghana\\Word lemma pos"
pathhke <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Hong Kong\\Word lemma pos"
pathide <- "D:\\Uni\\Korpora\\Original\\GloWbE\\India\\Word lemma pos"
pathire <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Ireland\\WordLemmaPos"
pathjme <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Jamaica\\Word lemma pos"
pathkne <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Kenya\\Word lemma pos"
pathmye <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Malaysia\\Word lemma pos"
pathnze <- "D:\\Uni\\Korpora\\Original\\GloWbE\\New Zealand\\Word lemma pos"
pathnge <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Nigeria\\Word lemma pos"
pathpke <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Pakistan\\Word lemma pos"
pathphe <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Philippines\\Word lemma pos"
pathsge <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Singapore\\Word lemma pos"
pathsae <- "D:\\Uni\\Korpora\\Original\\GloWbE\\South Africa\\Word lemma pos"
pathsle <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Sri Lanka\\Word lemma pos"
pathtze <- "D:\\Uni\\Korpora\\Original\\GloWbE\\Tanzania\\Word lemma pos"
pathusb <- "D:\\Uni\\Korpora\\Original\\GloWbE\\US_Blog\\Word lemma pos"
pathusg <- "D:\\Uni\\Korpora\\Original\\GloWbE\\US General\\Word lemma pos"
```


```{r}
# extract files
filesaue = list.files(path = pathaue, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesbge = list.files(path = pathbge, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filescae = list.files(path = pathcae, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesgbb = list.files(path = pathgbb, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesgbg = list.files(path = pathgbg, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesghe = list.files(path = pathghe, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
fileshke = list.files(path = pathhke, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
fileside = list.files(path = pathide, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesire = list.files(path = pathire, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesjme = list.files(path = pathjme, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
fileskne = list.files(path = pathkne, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesmye = list.files(path = pathmye, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesnze = list.files(path = pathnze, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesnge = list.files(path = pathnge, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filespke = list.files(path = pathpke, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesphe = list.files(path = pathphe, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filessge = list.files(path = pathsge, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filessae = list.files(path = pathsae, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filessle = list.files(path = pathsle, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filestze = list.files(path = pathtze, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesusb = list.files(path = pathusb, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
filesusg = list.files(path = pathusg, pattern = ".txt", all.files = T, full.names = T, recursive = T, ignore.case = T, include.dirs = T)
```


```{r}
# apply function
# apply function
glowbe_post(gfiles = filesaue) #(remove # to to activate)
glowbe_post(gfiles = filesbge) #(remove # to to activate)
glowbe_post(gfiles = filescae) #(remove # to to activate)
glowbe_post(gfiles = filesgbb) #(remove # to to activate)
glowbe_post(gfiles = filesgbg) #(remove # to to activate)
glowbe_post(gfiles = filesghe) #(remove # to to activate)
glowbe_post(gfiles = fileshke) #(remove # to to activate)
glowbe_post(gfiles = fileside) #(remove # to to activate)
glowbe_post(gfiles = filesire) #(remove # to to activate)
glowbe_post(gfiles = filesjme) #(remove # to to activate)
glowbe_post(gfiles = fileskne) #(remove # to to activate)
glowbe_post(gfiles = filesmye) #(remove # to to activate)
glowbe_post(gfiles = filesnze) #(remove # to to activate)
glowbe_post(gfiles = filesnge) #(remove # to to activate)
glowbe_post(gfiles = filespke) #(remove # to to activate)
glowbe_post(gfiles = filesphe) #(remove # to to activate)
glowbe_post(gfiles = filessge) #(remove # to to activate)
glowbe_post(gfiles = filessae) #(remove # to to activate)
glowbe_post(gfiles = filessle) #(remove # to to activate)
glowbe_post(gfiles = filestze) #(remove # to to activate)
glowbe_post(gfiles = filesusb) #(remove # to to activate)
glowbe_post(gfiles = filesusg) #(remove # to to activate)
```


```{r}
sessionInfo()
```


***

The END

***
