---
title: "X-much CXN in IrE - Part 4: Processing controll corpus"
author: "Anonymous"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  word_document
bibliography: bibliography.bib
link-citations: yes
---

This document focuses on the use of the X-much construction in Irish and Australian English. 

This part of the analysis performs the data processing of the control corpus.

# Session preparation

cleaning the workspace, loading packages, setting options, and defining paths.

```{r xmprep_2_01}
# load packages
library(tidyverse)
library(here)
library(readxl)
library(flextable)
library(quanteda)
library(tidytext)
# setting options
options(stringsAsFactors = F)
```

## Load adta

```{r load, message=F, warning=F}
ccps  <- base::readRDS(file = here::here("data", "ccorpus.rda")) %>%
  dplyr::mutate(docname = stringr::str_remove_all(paste0(file, "_", subfile), "/"))
xmd <- base::readRDS(file = here::here("data/kwics", "xmd.rda"))
xmdfiles <- unique(stringr::str_remove_all(xmd$docname, "\\..*"))
```

Remove files that are in the xmd data

```{r filter, message=F, warning=F}
ccps <- ccps %>%
  dplyr::filter(!docname %in% xmdfiles) %>%
  dplyr::mutate(variety = stringr::str_replace_all(docname, ".*_([a-z]{2,2})_.*", "\\1"))
# inspect
nrow(ccps)
```

## Extract adjectives, nouns, and verbs

```{r adj, message=F, warning=F}
# oz control
cctxt_au <- ccps %>%
  dplyr::filter(!docname %in% xmdfiles) %>%
  dplyr::mutate(text = paste0(variety, "_", text)) %>%
  dplyr::filter(stringr::str_detect(text, "^au")) %>%
  dplyr::pull(text)
# ie control
cctxt_ie <- ccps %>%
  dplyr::filter(!docname %in% xmdfiles) %>%
  dplyr::mutate(text = paste0(variety, "_", text)) %>%
  dplyr::filter(stringr::str_detect(text, "^ie")) %>%
  dplyr::pull(text)
# tok oz
tokcc_au <- quanteda::tokens(cctxt_au, what = "fastestword") %>%
  unlist() %>%
  sample(50000)
# tok ie
tokcc_ie <- quanteda::tokens(cctxt_ie, what = "fastestword") %>%
  unlist() %>%
  sample(50000)
# combine
tokcc <-c(tokcc_au, tokcc_ie)
# generate df
tkdf <- tokcc %>%
  as.data.frame() %>%
  dplyr::mutate(variety = c(rep("au", length(tokcc_au)), rep("ie", length(tokcc_ie)))) %>%
  dplyr::mutate(pos = stringr::str_remove_all(tokcc, ".*/"),
                word = tolower(stringr::str_remove_all(tokcc, "/.*"))) %>%
  dplyr::rename(full = colnames(.)[1]) %>%
  dplyr::mutate(pos = stringr::str_remove_all(pos, "[0-9].*|@.*|_.*"),
                pos = dplyr::case_when(pos == "vv" ~ pos,
                                       pos == "jj" ~ pos,
                                       pos == "nn" ~ pos,
                                       TRUE ~ "other"))
# inspect
head(tkdf)
```

add emotionality

```{r emo1, message=F, warning=F}
# sentiments
emo <- get_sentiments("nrc") %>%
  dplyr::filter(sentiment == "positive"| sentiment == "negative")
wordpos <- emo %>%
  dplyr::filter(sentiment == "positive") %>%
  dplyr::pull(word)
wordneg <- emo %>%
  dplyr::filter(sentiment == "negative") %>%
  dplyr::pull(word)
```


```{r emo1, message=F, warning=F}
cemodf <- tkdf %>%
  dplyr::mutate(emo = dplyr::case_when(word %in% wordpos ~ "positive",
                                       word %in% wordneg ~ "negative",
                                       TRUE ~ "nonemotional")) %>%
  dplyr::mutate(control = "control")
# inspect
head(cemodf)
```

# Differences in emotionality

```{r}
xmdtb <- xmd %>%
  dplyr::select(variety, x, type, emo) %>%
  dplyr::rename(word = x,
                pos = type) %>%
  dplyr::mutate(control = "test",
                emocat = ifelse(emo == "nonemotional", emo, "emotional"),
                pos = dplyr::case_when(pos == "vv" ~ pos,
                                       pos == "jj" ~ pos,
                                       pos == "nn" ~ pos,
                                       TRUE ~ "other"))
# inspect
head(xmdtb)
```

```{r}
cemotb <- cemodf %>%
  dplyr::mutate(emocat = ifelse(emo == "nonemotional", emo, "emotional"),
                variety = ifelse(variety == "au", "Australia", "Ireland")) %>%
  dplyr::select(-full) %>%
  dplyr::relocate(variety, word)
# inspect
head(cemotb)
```


```{r}
xmdj <- rbind(xmdtb,cemotb) 
# inspect
head(xmdj)
```

# Statistics

# Emotionality

## X2:  Emotionality vs Control

RQ: Does the emotionality of adjectives in the test data (adjectives in the x-much construction) differ from the emotionality in the control data?

```{r}
emo_tb1 <- xmdj %>%
  dplyr::filter(pos == "jj") %>%
  dplyr::group_by(emocat, control) %>%
  dplyr::summarise(freq = n())
# inspect
head(emo_tb1)
```

Perform x2-test

```{r}
emo_x2 <- emo_tb1 %>%
  tidyr::spread(control, freq) %>%
  as.matrix()
rn<- emo_x2[,1]
emo_x2 = emo_x2[, 2:3]
emo_x2 <- apply(emo_x2, 2, as.numeric)
rownames(emo_x2) <- rn
emo_x2
# perform x2 test
chisq.test(emo_x2)
```

The adjectives in the x-much construction are not more emotional compared to the adjectives in the controll data.

## CFA: Emotions vs Control

RQ: Do the adjectives in the test data (adjectives in the x-much construction) differ from the emotionality in the control data when differentiating between positive, nonemotional, and negative emotional?

```{r}
emo_tb2 <- xmdj %>%
  dplyr::filter(pos == "jj") %>%
  dplyr::group_by(emo, control) %>%
  dplyr::summarise(freq = n())
# inspect
emo_tb2
```

Perform CFA

```{r}
library(cfa)
configs <- emo_tb2 %>%
  dplyr::select(control, emo)
counts = emo_tb2$freq
cfa::cfa(configs, counts)
```

The adjectives in the x-much construction are significantly more negative-emotional compared to the adjectives in the control data.

##  CFA: Emotions vs Variety

```{r}
emo_tb3 <- xmdj %>%
  dplyr::filter(pos == "jj") %>%
  dplyr::group_by(emo, variety) %>%
  dplyr::summarise(freq = n())
# inspect
emo_tb3
```

```{r}
configs <- emo_tb3 %>%
  dplyr::select(variety, emo)
counts = emo_tb3$freq
cfa::cfa(configs, counts)
```

There are significantly more negative adjectives in the Australian data after corrections compared to any other configuation.


# Word Prefernces

## CFA: Pos attracted to x-much vs control

RQ: are certain pos significantly more common in the x-much construction compared to the control data?


```{r}
poscat_tb <- xmdj %>%
  dplyr::group_by(pos, control) %>%
  dplyr::summarise(freq = n())
configs <- poscat_tb %>%
  dplyr::select(control, pos)
counts = poscat_tb$freq
cfa::cfa(configs, counts)
```

According to teh CFA, jj, nn, and vv are significantly overrepresented while other pos are significantly underrepresented.

## CFA: Pos attracted to x-much vs variety


RQ: are certain pos significantly more common in the x-much construction by variety?

```{r}
poscat_tb <- xmdj %>%
  dplyr::group_by(pos, variety) %>%
  dplyr::summarise(freq = n())
configs <- poscat_tb %>%
  dplyr::select(variety, pos)
counts = poscat_tb$freq
cfa::cfa(configs, counts)
```

In Australian English jj, vv, nn, **and other** are significantly over-represented. In Irish English only nn and jj.

**x-much has expanded/less restrictive than in IrE**


## CFA: words overrepresented in x-much

```{r}
wordcat_tb <- xmdj %>%
  dplyr::group_by(word, control) %>%
  dplyr::summarise(freq = n())
configs <- wordcat_tb %>%
  dplyr::select(control, word)
counts = wordcat_tb$freq
cfa::cfa(configs, counts)
```




## Save data to disc{-}

```{r}
base::saveRDS(emodf, file = here::here("tables", "cc_emo.rda"))
```


# Outro

```{r}
sessionInfo()
```





